
%
% Acknowledgements
%

% Guideline
% An optional section which acknowledges the contributions of other researchers or organisations which have made facilities available to the researcher. No page numbering is ascribed to the Acknowledgements section

% title
\chapter*{Acknowledgements}
\label{acknowledgements}
% ensure it appears in the doc
\addcontentsline{toc}{chapter}{Acknowledgements}
I could not have completed this work without the help and assistance of a large number of people and organisations.

%
% Professionally
%
\subsubsection*{Professionally, I would like to thank:} 
% tim
\emph{Tim Hendtlass} for your wisdom, enthusiasm, encouragement and seemingly unbounded patience and support through times easy and difficult. Your hands off approach has made me both more mature as a person, and more confident as a researcher. I am fortunate to have counted you as mentor and friend these past years.
% swinburne
The \emph{Faculty of Information and Communication Technologies, Swinburne University of Technology} for providing me with candidature and support throughout my PhD.
% government
The \emph{Australian Government}, for providing me with the financial assistance of an Australian Postgraduate Award (APA) during my candidature.


%
% NOTE: tim - not having fun, not doing it right, about enjoyment, excitement, etc.
%


%
% Personally
%
\subsubsection*{Personally, I would like to thank:}
% dan
\emph{Daniel Angus} for his friendship, discussions, rants, debates, collaboration, competition, and ultimately for defining my PhD experience.
% colleague's
My officemates, past and present: \emph{Irene Moser}, \emph{Nigel Fisher}, \emph{Gerard Murray}, \emph{Clinton Woodward}, \emph{Hans Pierrot}, \emph{James Montgomery}, and \emph{Ben Prime} for providing much needed fun, drinking partners, and most importantly anchors for my sanity.
\emph{Friends and Housemates} for the good times, bad times, and the much needed balance you brought to my life.
% ying
\emph{Ying Liu} for your intimate understanding and patience for me whilst writing up my thesis.
% Family
Finally, my parents \emph{Lynette Brownlee} and \emph{Derek Brownlee}, and my sister \emph{Laura Brownlee} for your unconditional continued support of my decision to quit a perfectly good job and pursue my research interests.

% EOF